Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach

### 2019-01-12 Manuscript.Rmd
More detailed abstract

  Valuation effects are consistently observed in medial prefrontal and posterior cingulate cortex (mPFC and PCC). The spatial extent of these effects is mostly indistinguishable from the default mode network (DMN) in existing meta-analyses. However, little is known about how valuation effects fit within the broader functional architecture of mPFC and PCC, or whether that architecture is consistent or idiosyncratic across individuals. Here we complement a meta-analysis with fMRI-based graph theoretic approaches to subdivide mPFC and PCC at the single-subject level. Our results suggest the functional topography of mPFC has substantial variability across individuals. This highlights the potential usefulness of estimating brain effects at the individual level in this region, and points to limitations of aggregative methods such as coordinate-based meta-analysis in determining whether valuation and DMN effects emerge from common or separable brain systems. Our approach shows promise in addressing this issue through future manipulations of valuation.

>>>>

Valuation effects are consistently observed in medial prefrontal and posterior cingulate cortex (mPFC and PCC). The spatial extent of these effects is mostly indistinguishable from the default mode network (DMN) in existing meta-analyses. However, little is known about how valuation effects fit within the broader functional architecture of mPFC and PCC, or whether that architecture is consistent or idiosyncratic across individuals. Here we complement a meta-analysis with fMRI-based graph-theoretic approaches on functional connectivity to subdivide mPFC and PCC at the single-subject level. Our results show that valuation and DMN effects had substantial meta-analytic overlap in mPFC and PCC. However, the regions of group-level overlap could be reliably segregated into two communities in every individual (n = 100). This subdivision had low inter-individual alignment despite sharing qualitative topological features similar to those found by Braga and Buckner (2017). Using measures that correct for chance clustering, we found that (1) different graph-theoretic clustering algorithms strongly agree in this partition; (2) each individual’s network arrangement was more similar across their own fMRI sessions than across subjects; (3) this idiosyncrasy was higher for mPFC; and (4) the idiosyncratic functional topography was not merely the result of measurement noise. We also show the advantages of graph-theoretic approaches over traditional seed-based functional connectivity for single-subject analysis. This work highlights the potential usefulness of estimating brain effects at the individual level in these regions, and points to limitations of aggregative methods in determining the degree of neural overlap between DMN and valuation. Future behavioral manipulations will examine whether the two sub-networks are better understood as subdivisions of the DMN or as segregating DMN from non-DMN regions.
